---
title: "07_biomarkerFilter_onlyANOVA"
output: html_document
date: "2026-01-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load data

```{r}
load("../data/biomarker_input/anova_features_WITH_duplicates.RData")
load("../data/biomarker_input/PCA_features_WITHduplicates.RData")
load("../data/biomarker_input/NMF_features_WITH_dupes.RData")
load("../data/biomarker_input/NMF_feature_scores_WITH_dupes.RData")
```

# Load packages

```{r}
packages <- c("tidyverse", "dplyr", "ggplot2", "ggpattern", "plotly", 
              "RColorBrewer", "ggsci", "ggpubr", "pheatmap", "grafify", "patchwork",
              "openxlsx", "data.table", "grafify", "factoextra", "htmltools", "colorspace",
              "VennDiagram", "gridExtra")

installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages])
}



biocpackage <- "ComplexHeatmap"
if((biocpackage %in% rownames(installed.packages())) == FALSE) {
  BiocManager::install("ComplexHeatmap")
}
invisible(lapply(c(packages,biocpackage), library, character.only = TRUE))
library(ComplexHeatmap)

source("../code/reproducibility_functions.R")
```

# Only include M+H, M-H, M+e, or M-e features

```{r}
feature.filt <- function(data, comp){
  # data = results from analysis of one lab
  # comp = metabolite meta-data of one lab
  
    # add feature name to data
    
    comp <- rownames_to_column(as.data.frame(comp), var = "feature")
    
    if(any(grepl("feature", colnames(data))) == FALSE){
      
      new.data <- rownames_to_column(as.data.frame(data), var = "feature")
      
    }else{
      
      new.data <- data
      
    }
    
    new.data <- merge(new.data, subset(comp, select = c(feature, Name)))
  
    # only include M+H, M-H, M+e, and M-e features
    
    include <- c("(M+H)", "(M-H)", "(M+e)", "(M-e)")

    adducts <- gsub("^\\S* ", "", new.data$Name)
    
    fin.data <- new.data[which(adducts %in% include),]
    
    # remove name column
    
    fin.data <- subset(fin.data, select = -c(Name))
    
    # if feature wasn't originally part of data, remove
    
    if(any(grepl("feature", colnames(data))) == FALSE){
      
      row.names(fin.data) <- fin.data$feature
      
      fin.data <- subset(fin.data, select = -c(feature))
      
    }
    
    return(fin.data)
    
}

# Apply to NMF data

nmf.feat.filt <-list()
length(nmf.feat.filt) <- length(scores)
names(nmf.feat.filt) <- names(scores)

for(lab in 1:length(scores)){
  
  nmf.feat.filt[[lab]] <- feature.filt(scores[[lab]], comp.found.na.filter[[lab]])
  
}

# Apply to ANOVA data
anova.feat.filt <-list()
length(anova.feat.filt) <- length(anova.results)
names(anova.feat.filt) <- names(anova.results)

for(lab in 1:length(scores)){
  
  anova.feat.filt[[lab]] <- feature.filt(anova.results[[lab]], comp.found.na.filter[[lab]])
  
}
```

# New NMF features filter
feature with highest score per MF & score in top 50%

```{r}
nmf.filter.score <- function(data, comp, percentile = 0.25){
  # data = list of feature scores per lab
  # comp = metabolite meta-data
  # percentile = percentile cutoff for top scores
  
  # Add MF and name to scores
  
  score.mf <- rownames_to_column(as.data.frame(data), var = "feature")
  comp.mf <- rownames_to_column(comp, var = "feature")
  
  score.mf <- merge(score.mf, subset(comp.mf, select = c(feature, Name, MF)), by = "feature")
  score.mf$Name <- gsub(" .*", "", score.mf$Name)
  
  # Find scores in top percentile
  
  perc <- 1 - percentile
  
  cutoff <- quantile(score.mf$data, perc)
  
  score.top <- score.mf[which(score.mf$data > cutoff),]
  
  # Find feature with highest score per MF
      
      mfs <- unique(score.top$MF)
      
      feats.fin <- data.frame(matrix(nrow = length(mfs), ncol = 4))
      colnames(feats.fin) <- colnames(score.mf)
      
      for(mf in 1:nrow(feats.fin)){
        
        feats <- score.top[which(score.top$MF == mfs[mf]),]
        
        feats.fin[mf,] <- feats[which(feats$data == max(feats$data)),]
        
      }
  return(feats.fin)
}

nmf.feats <-list()
length(nmf.feats) <- length(nmf.feat.filt)
names(nmf.feats) <- names(nmf.feat.filt)

for(lab in 1:length(scores)){
  
  nmf.feats[[lab]] <- nmf.filter.score(nmf.feat.filt[[lab]], comp.found.na.filter[[lab]], percentile = 0.5)
  
}

```

# NMF features filter
feature with highest contribution to a single rank in NMF of features that contribute >66% to a given rank

```{r}
nmf.filter <- function(data, comp){
  
    # find features that contribute >66% to a given rank
      
  feats.66 <- apply(data, 2, function(x) any(x > 0.66))
      feats.66 <- data[,which(feats.66 == TRUE)]
     
      # make a separate table with the feature, rank, and MF
      
      feats.info <- data.frame(matrix(nrow = ncol(feats.66), ncol = 3))
      colnames(feats.info) <- c("feature", "rank", "rank.perc")
      feats.info$feature <- colnames(feats.66)
       
      # add MFs
      
      comp <- as.data.frame(comp)
      comp.feats <- rownames_to_column(comp, var = "feature")
      feats.info <- merge(feats.info, subset(comp.feats, select = c(feature, MF)), by = "feature")
     
      # add rank and percent
      
      for(feat.filter in 1:nrow(feats.info)){
        for(feat.orig in 1:ncol(data)){
          
          if(feats.info$feature[feat.filter] == colnames(data)[feat.orig]){
            
           feats.info$rank.perc[feat.filter] <- max(data[,feat.orig]) 
           
           feats.info$rank[feat.filter] <- names(which(data[,feat.orig] == max(data[,feat.orig])))
            
          }  
        }
      }
   
    # find the feature in each MF that contributes to the rank the most
      
      mfs <- unique(feats.info$MF)
      
      feats.fin <- data.frame(matrix(nrow = length(mfs), ncol = 4))
      colnames(feats.fin) <- colnames(feats.info)
      
      for(mf in 1:nrow(feats.fin)){
        
        feats <- feats.info[which(feats.info$MF == mfs[mf]),]
        
        feats.fin[mf,] <- feats[which(feats$rank.perc == max(feats$rank.perc)),]
        
      }
      
  return(feats.fin)    
      
}

#nmf.feats <-list()
#length(nmf.feats) <- length(h)
#names(nmf.feats) <- names(h)

#for(lab in 1:length(h)){
  
 # nmf.feats[[lab]] <- nmf.filter(h[[lab]], comp.found.na.filter[[lab]])
  
#}

```

# Subset abundance data for features filtered based on: NMF

```{r}
# Load data for heatmap abundances
load("../data/duplicate_input/imputed_data.RData")

# Unlist MFs to get master list
  mf.unlist <- sapply(nmf.feats, "[[", "MF")
  mf.unlist <- unlist(mf.unlist)
  mf.unique <- unique(mf.unlist)
  
  nmf.mf.list <- list()
  length(nmf.mf.list) <- length(mf.unique)
  names(nmf.mf.list) <- mf.unique

# Add abundance data

  for(mf in 1:length(nmf.mf.list)){
    
    if(mf.unique[mf] %in% nmf.feats$Lab1$MF &
       mf.unique[mf] %in% nmf.feats$Lab2$MF &
       mf.unique[mf] %in% nmf.feats$Lab3$MF &
       mf.unique[mf] %in% nmf.feats$Lab4$MF &
       mf.unique[mf] %in% nmf.feats$Lab5$MF &
       mf.unique[mf] %in% nmf.feats$Lab6$MF &
       mf.unique[mf] %in% nmf.feats$Lab7$MF &
       mf.unique[mf] %in% nmf.feats$Lab8$MF){
      
      # add empty data frame to list
      
      nmf.mf.list[[mf]] <- data.frame(matrix(nrow = 8, ncol = nrow(norm.impute$Lab1)))
      colnames(nmf.mf.list[[mf]]) <- row.names(norm.impute$Lab1)
      row.names(nmf.mf.list[[mf]]) <- names(norm.impute)
      
      # add abundance data for the feature from the filtered feature list for each MF for each lab
      
      nmf.mf.list[[mf]][1,] <- norm.impute$Lab1[,which(row.names(comp.found.na.filter$Lab1) == nmf.feats$Lab1$feature[which(nmf.feats$Lab1$MF == mf.unique[mf])])]
      nmf.mf.list[[mf]][2,] <- norm.impute$Lab2[,which(row.names(comp.found.na.filter$Lab2) == nmf.feats$Lab2$feature[which(nmf.feats$Lab2$MF == mf.unique[mf])])]
      nmf.mf.list[[mf]][3,] <- norm.impute$Lab3[,which(row.names(comp.found.na.filter$Lab3) == nmf.feats$Lab3$feature[which(nmf.feats$Lab3$MF == mf.unique[mf])])]
      nmf.mf.list[[mf]][4,] <- norm.impute$Lab4[,which(row.names(comp.found.na.filter$Lab4) == nmf.feats$Lab4$feature[which(nmf.feats$Lab4$MF == mf.unique[mf])])]
      nmf.mf.list[[mf]][5,] <- norm.impute$Lab5[,which(row.names(comp.found.na.filter$Lab5) == nmf.feats$Lab5$feature[which(nmf.feats$Lab5$MF == mf.unique[mf])])]
      nmf.mf.list[[mf]][6,] <- norm.impute$Lab6[,which(row.names(comp.found.na.filter$Lab6) == nmf.feats$Lab6$feature[which(nmf.feats$Lab6$MF == mf.unique[mf])])]
      nmf.mf.list[[mf]][7,] <- norm.impute$Lab7[,which(row.names(comp.found.na.filter$Lab7) == nmf.feats$Lab7$feature[which(nmf.feats$Lab7$MF == mf.unique[mf])])]
      nmf.mf.list[[mf]][8,] <- norm.impute$Lab8[,which(row.names(comp.found.na.filter$Lab8) == nmf.feats$Lab8$feature[which(nmf.feats$Lab8$MF == mf.unique[mf])])]
      
    }
  }

# Remove NULLs
  
  nmf.mf.list <- nmf.mf.list[!sapply(nmf.mf.list, is.null)]

```

# Take the average abundance of each group

```{r}
avg.abund <- function(data, meta.data){
  # data = filtered list with metabolite abundances for each MF (labs in rows, samples in columns)
  # meta.data = meta data with group information
  
  # Prep results table
  avg.results <- data.frame(matrix(nrow = nrow(data), ncol = length(unique(meta.data$group))))
  colnames(avg.results) <- unique(meta.data$group)
  row.names(avg.results) <- row.names(data)
  
  # Calculate averages and add to results table
  
  for(group in 1:ncol(avg.results)){
    
    data.in <- data[,which(meta.data$group == colnames(avg.results)[group])]
    
    avg.results[,group] <- rowMeans(data.in)  
    
  }
  
  return(avg.results)
  
}

nmf.mf.avg <- lapply(nmf.mf.list, function(x) avg.abund(x, meta.data))
```

# Remove peptides from the metabolites of interest: NMF

```{r}
# List MFs that correspond to putative peptide IDs

peptide.mf <- c("C10H16N2O4", "C10H18N2O5", "C10H19N3O4", "C10H20N2O3", "C11H20N2O3", "C11H20N2O5", "C11H22N2O3", "C11H23N5O3", "C12H16N2O3", "C12H25N3O3", "C13H18N2O4", "C13H23N3O4", "C14H20N2O3", "C14H20N2O4", "C15H22N2O3", "C15H22N2O4", "C16H21N3O3", "C8H14N2O4", "C8H16N2O3", "C9H17N3O4", "C9H17N3O4", "C9H18N2O3", "C9H18N2O4", "C17H31N3O4", "C16H29N3O4")

# Remove peptides

index <- which(names(nmf.mf.list) %in% peptide.mf)  
  
nmf.mf.list.nopep <- nmf.mf.list[-index]
```

# Order meta-data for all heatmaps

```{r}
meta.data$group <- factor(meta.data$group, levels = c("a", "b", "c", "ab", "ac", "bc", "abc"))

meta.data.avg <- meta.data[which(duplicated(meta.data$group) == F),]
meta.data.avg$group <- factor(meta.data.avg$group, levels = c("a", "b", "c", "ab", "ac", "bc", "abc"))
```

# Make heatmaps of abundance data: NMF - all abundances

```{r}
nmf.scaled.data <- lapply(nmf.mf.list, function(y) apply(y, 1, function(x) log2(x/(mean(x)))))

colors <- colorRampPalette(c("white", "lightblue", "darkblue"))(50)

plots.nmf.all <- mapply(function(x,y) (pheatmap::pheatmap(t(x[order(meta.data$group),]), cluster_rows = F, cluster_cols = F, main = y, 
                                             color = colors,
                                             show_rownames = T,
                                             show_colnames = T,
                                             fontsize = 15,
                                             silent = TRUE)$gtable), 
                                           nmf.scaled.data, 
                                           names(nmf.scaled.data),
                                           SIMPLIFY = FALSE)

pdf(file = "../outputs/heatmap_NMF_feats.pdf", width = 24, height = 16)

grid.arrange(grobs = plots.nmf.all, ncol = 4)

dev.off()

mapply(function(x,y) (pheatmap::pheatmap(t(x[order(meta.data$group),]), cluster_rows = F, cluster_cols = F, main = y, 
                                             color = colors,
                                             show_rownames = F,
                                             show_colnames = F)), nmf.scaled.data, names(nmf.scaled.data))

pheatmap::pheatmap(t(nmf.scaled.data$C4H7NO4[order(meta.data$group),]), cluster_rows = F, cluster_cols = F, color = colors, show_rownames = T, show_colnames = T)
```

# Make heatmaps of abundance data: NMF - averaged abundances

```{r}
nmf.scaled.data.avg <- lapply(nmf.mf.avg, function(y) apply(y, 1, function(x) log2(x/(mean(x)))))

# get putative names for MFs
mf.names <- gsub(" .*", "", comp.found.na.filter$Lab1$Name[which(comp.found.na.filter$Lab1$MF %in% names(nmf.scaled.data.avg))])

names(mf.names) <- comp.found.na.filter$Lab1$MF[which(comp.found.na.filter$Lab1$MF %in% names(nmf.scaled.data.avg))]

mf.names <- mf.names[which(duplicated(names(mf.names)) == F)]

  ## fix that long one

mf.names[c("C16H29N3O4")] <- "CHEBI:118698"

# set colors
colors <- colorRampPalette(c("white", "lightblue", "darkblue"))(50)

# plot
plots.nmf.avg <- mapply(function(x,y,z) (pheatmap::pheatmap(t(x[order(meta.data.avg$group),]), cluster_rows = F, cluster_cols = F, main = paste0(y, " - ", z), 
                                             color = colors,
                                             show_rownames = T,
                                             show_colnames = T,
                                             fontsize = 15,
                                             fontsize_col = 20,
                                             fontsize_row = 20,
                                             silent = TRUE)$gtable), 
                                           nmf.scaled.data.avg, 
                                           names(nmf.scaled.data.avg),
                                           mf.names,
                                           SIMPLIFY = FALSE)

pdf(file = "../outputs/heatmap_NMF_feats_averaged.pdf", width = 24, height = 16)

grid.arrange(grobs = plots.nmf.avg, ncol = 4)

dev.off()

mapply(function(x,y,z) (pheatmap::pheatmap(t(x[order(meta.data.avg$group),]), cluster_rows = F, cluster_cols = F, main = paste0(y, " - ", z), 
                                             color = colors,
                                             show_rownames = F,
                                             show_colnames = F)), 
       nmf.scaled.data.avg, 
       names(nmf.scaled.data.avg),
       mf.names)
```

# ANOVA features filter
feature with lowest p-value for a given MF

```{r}
anova.filter <- function(data, p.cutoff = 0.05){
  # data = anova results 
  # p.cutuff = adjusted p-value threshold
 
  # Find features with lowest p-value under 0.05
   
    ## only include features with adj p-value < 0.05
  
    data.filter <- data[which(data$adj.p < p.cutoff),]
    
    ## set up results table
    
    mfs <- unique(data.filter$MF)
    
    results <- as.data.frame(matrix(nrow = length(mfs), ncol = ncol(data.filter)))
    colnames(results) <- colnames(data.filter)
    
    ## find the lowest p-value
    
    for(mf in 1:length(mfs)){
      
      feats <- data.filter[which(data.filter$MF == mfs[mf]),]
        
      low.p <- feats[which(feats$adj.p == min(feats$adj.p)),]
      
      results[mf,] <- low.p
    
    }
  
  return(results)
  
}

anova.feats <- lapply(anova.feat.filt, function(x) anova.filter(x, 0.05))
```

# Subset abundance data for features filtered based on: ANOVA

```{r}
# Load data for heatmap abundances
load("../data/duplicate_input/imputed_data.RData")

# Unlist MFs to get master list
  mf.unlist <- sapply(anova.feats, "[[", "MF")
  mf.unlist <- unlist(mf.unlist)
  mf.unique <- unique(mf.unlist)
  
  anova.mf.list <- list()
  length(anova.mf.list) <- length(mf.unique)
  names(anova.mf.list) <- mf.unique

# Add abundance data

  for(mf in 1:length(anova.mf.list)){
    
    if(mf.unique[mf] %in% anova.feats$Lab1$MF &
       mf.unique[mf] %in% anova.feats$Lab2$MF &
       mf.unique[mf] %in% anova.feats$Lab3$MF &
       mf.unique[mf] %in% anova.feats$Lab4$MF &
       mf.unique[mf] %in% anova.feats$Lab5$MF &
       mf.unique[mf] %in% anova.feats$Lab6$MF &
       mf.unique[mf] %in% anova.feats$Lab7$MF &
       mf.unique[mf] %in% anova.feats$Lab8$MF){
      
      # add empty data frame to list
      
      anova.mf.list[[mf]] <- data.frame(matrix(nrow = 8, ncol = nrow(norm.impute$Lab1)))
      colnames(anova.mf.list[[mf]]) <- row.names(norm.impute$Lab1)
      row.names(anova.mf.list[[mf]]) <- names(norm.impute)
      
      # add abundance data for the feature from the filtered feature list for each MF for each lab
      
      anova.mf.list[[mf]][1,] <- norm.impute$Lab1[,which(row.names(comp.found.na.filter$Lab1) == anova.feats$Lab1$feature[which(anova.feats$Lab1$MF == mf.unique[mf])])]
      anova.mf.list[[mf]][2,] <- norm.impute$Lab2[,which(row.names(comp.found.na.filter$Lab2) == anova.feats$Lab2$feature[which(anova.feats$Lab2$MF == mf.unique[mf])])]
      anova.mf.list[[mf]][3,] <- norm.impute$Lab3[,which(row.names(comp.found.na.filter$Lab3) == anova.feats$Lab3$feature[which(anova.feats$Lab3$MF == mf.unique[mf])])]
      anova.mf.list[[mf]][4,] <- norm.impute$Lab4[,which(row.names(comp.found.na.filter$Lab4) == anova.feats$Lab4$feature[which(anova.feats$Lab4$MF == mf.unique[mf])])]
      anova.mf.list[[mf]][5,] <- norm.impute$Lab5[,which(row.names(comp.found.na.filter$Lab5) == anova.feats$Lab5$feature[which(anova.feats$Lab5$MF == mf.unique[mf])])]
      anova.mf.list[[mf]][6,] <- norm.impute$Lab6[,which(row.names(comp.found.na.filter$Lab6) == anova.feats$Lab6$feature[which(anova.feats$Lab6$MF == mf.unique[mf])])]
      anova.mf.list[[mf]][7,] <- norm.impute$Lab7[,which(row.names(comp.found.na.filter$Lab7) == anova.feats$Lab7$feature[which(anova.feats$Lab7$MF == mf.unique[mf])])]
      anova.mf.list[[mf]][8,] <- norm.impute$Lab8[,which(row.names(comp.found.na.filter$Lab8) == anova.feats$Lab8$feature[which(anova.feats$Lab8$MF == mf.unique[mf])])]
      
    }
  }

# Remove NULLs
  
  anova.mf.list <- anova.mf.list[!sapply(anova.mf.list, is.null)]

```

# Remove peptides from the metabolites of interest: ANOVA

```{r}
# List MFs that correspond to putative peptide IDs

peptide.mf <- c("C10H16N2O4", "C10H18N2O5", "C10H19N3O4", "C10H20N2O3", "C11H20N2O3", "C11H20N2O5", "C11H22N2O3", "C11H23N5O3", "C12H16N2O3", "C12H25N3O3", "C13H18N2O4", "C13H23N3O4", "C14H20N2O3", "C14H20N2O4", "C15H22N2O3", "C15H22N2O4", "C16H21N3O3", "C8H14N2O4", "C8H16N2O3", "C9H17N3O4", "C9H17N3O4", "C9H18N2O3", "C9H18N2O4", "C17H31N3O4", "C16H29N3O4")

# Remove peptides

index <- which(names(anova.mf.list) %in% peptide.mf)  
  
anova.mf.list.nopep <- anova.mf.list[-index]
```

# Take the average abundance for each group: ANOVA

```{r}
anova.mf.avg <- lapply(anova.mf.list, function(x) avg.abund(x, meta.data))
```

# Make heatmaps of abundance data: ANOVA - all abundances

```{r}
anova.scaled.data <- lapply(anova.mf.list, function(y) apply(y, 1, function(x) log2(x/(mean(x)))))

colors <- colorRampPalette(c("white", "lightblue", "darkblue"))(50)

plots.anova.all <- mapply(function(x,y) (pheatmap::pheatmap(t(x[order(meta.data$group),]), cluster_rows = F, cluster_cols = F, main = y, 
                                             color = colors,
                                             show_rownames = T,
                                             show_colnames = T,
                                             silent = TRUE)$gtable), 
                                           anova.scaled.data, 
                                           names(anova.scaled.data),
                                           SIMPLIFY = FALSE)

pdf(file = "../outputs/heatmap_anova_feats.pdf", width = 24, height = 16)

grid.arrange(grobs = plots.anova.all, ncol = 6)

dev.off()

mapply(function(x,y) (pheatmap::pheatmap(t(x[order(meta.data$group),]), cluster_rows = F, cluster_cols = F, main = y, 
                                             color = colors,
                                             show_rownames = F,
                                             show_colnames = F)), anova.scaled.data, names(anova.scaled.data))
```

# Make heatmaps of abundance data: ANOVA - averaged abundances

```{r}
anova.scaled.data.avg <- lapply(anova.mf.avg, function(y) apply(y, 1, function(x) log2(x/(mean(x)))))

colors <- colorRampPalette(c("white", "lightblue", "darkblue"))(50)

plots.anova.all <- mapply(function(x,y) (pheatmap::pheatmap(t(x[order(meta.data.avg$group),]), cluster_rows = F, cluster_cols = F, main = y, 
                                             color = colors,
                                             show_rownames = T,
                                             show_colnames = T,
                                             silent = TRUE)$gtable), 
                                           anova.scaled.data.avg, 
                                           names(anova.scaled.data.avg),
                                           SIMPLIFY = FALSE)

pdf(file = "../outputs/heatmap_anova_feats_averaged.pdf",width = 24, height = 16)

grid.arrange(grobs = plots.anova.all, ncol = 6)

dev.off()

mapply(function(x,y) (pheatmap::pheatmap(t(x[order(meta.data.avg$group),]), cluster_rows = F, cluster_cols = F, main = y, 
                                             color = colors,
                                             show_rownames = F,
                                             show_colnames = F)), anova.scaled.data.avg, names(anova.scaled.data.avg))
```

# Compare NMF and ANOVA features

```{r}
venn.diagram(list(names(nmf.scaled.data), names(anova.scaled.data)),
             category.names = c("NMF", "ANOVA"),
             filename = "../outputs/NMFvANOVA_feats_venn.png", 
             output = TRUE)

```


# PCA features filter
feature with highest loading in PC1

```{r}
pca.filter <- function(data, comp){
  
  # add MF to data
  
    comp <- as.data.frame(comp)
    comp <- rownames_to_column(comp, var = "feature")
  
    pca.feats <- data.frame(matrix(nrow = nrow(data), ncol = 2))
    colnames(pca.feats) <- c("feature", "loading")
    pca.feats$feature <- rownames(data)
    pca.feats$loading <- data$PC1
    
    pca.feats <- merge(pca.feats, subset(comp, select = c(feature, MF)), by = "feature")
  
  # find the feature with the highest absolute value loading for each MF
      
      mfs <- unique(pca.feats$MF)
      
      feats.fin <- data.frame(matrix(nrow = length(mfs), ncol = 3))
      colnames(feats.fin) <- colnames(pca.feats)
      
      for(mf in 1:nrow(feats.fin)){
        
        feats <- pca.feats[which(pca.feats$MF == mfs[mf]),]
        
        feats.fin[mf,] <- feats[which(abs(feats$loading) == max(abs(feats$loading))),]
        
      }
      
  return(feats.fin)
      
}

pca.feats <-list()
length(pca.feats) <- length(norm.group.load)
names(pca.feats) <- names(norm.group.load)

for(lab in 1:length(norm.group.load)){
  
  pca.feats[[lab]] <- pca.filter(norm.group.load[[lab]], comp.found.na.filter[[lab]])
  
}
```


# All overlapped features filter
features with retained names in all 8 labs

```{r}
# Load data with retained name info
load("../data/reproducibility_input/duplicated_filtered_data.RData")

overlap.feats <- overlap.metabs(comp.std, "retain.name")
```

# Subset abundance data for features filtered based on: overlaps

```{r}
# Prep list

length(overlap.mf.list) <- length(overlap.feats)
names(overlap.mf.list) <- overlap.feats


# Add abundance data

  for(feat in 1:length(overlap.feats)){
      
      # add empty data frame to list
      
      overlap.mf.list[[feat]] <- data.frame(matrix(nrow = 8, ncol = nrow(norm.std$Lab1)))
      colnames(overlap.mf.list[[feat]]) <- row.names(norm.std$Lab1)
      row.names(overlap.mf.list[[feat]]) <- names(norm.std)
      
      # add abundance data for the feature from the filtered feature list for each MF for each lab
      
      overlap.mf.list[[feat]][1,] <- norm.std$Lab1[,which(comp.std$Lab1$retain.name == overlap.feats[feat])]
      overlap.mf.list[[feat]][2,] <- norm.std$Lab2[,which(comp.std$Lab2$retain.name == overlap.feats[feat])]
      overlap.mf.list[[feat]][3,] <- norm.std$Lab3[,which(comp.std$Lab3$retain.name == overlap.feats[feat])]
      overlap.mf.list[[feat]][4,] <- norm.std$Lab4[,which(comp.std$Lab4$retain.name == overlap.feats[feat])]
      overlap.mf.list[[feat]][5,] <- norm.std$Lab5[,which(comp.std$Lab5$retain.name == overlap.feats[feat])]
      overlap.mf.list[[feat]][6,] <- norm.std$Lab6[,which(comp.std$Lab6$retain.name == overlap.feats[feat])]
      overlap.mf.list[[feat]][7,] <- norm.std$Lab7[,which(comp.std$Lab7$retain.name == overlap.feats[feat])]
      overlap.mf.list[[feat]][8,] <- norm.std$Lab8[,which(comp.std$Lab8$retain.name == overlap.feats[feat])]
      
  }

# Remove NULLs
  
  overlap.mf.list <- overlap.mf.list[!sapply(overlap.mf.list, is.null)]

```

# Remove peptides from the metabolites of interest: overlaps

```{r}
# List MFs that correspond to putative peptide IDs

peptide.mf <- c("C10H16N2O4", "C10H18N2O5", "C10H19N3O4", "C10H20N2O3", "C11H20N2O3", "C11H20N2O5", "C11H22N2O3", "C11H23N5O3", "C12H16N2O3", "C12H25N3O3", "C13H18N2O4", "C13H23N3O4", "C14H20N2O3", "C14H20N2O4", "C15H22N2O3", "C15H22N2O4", "C16H21N3O3", "C8H14N2O4", "C8H16N2O3", "C9H17N3O4", "C9H17N3O4", "C9H18N2O3", "C9H18N2O4", "C17H31N3O4", "C16H29N3O4")

# Remove peptides

index <- which(names(overlap.mf.list) %in% peptide.mf)  
  
overlap.mf.list.nopep <- overlap.mf.list[-index]
```

# Take the average abundance for each group: overlaps

```{r}
overlap.mf.avg <- lapply(overlap.mf.list, function(x) avg.abund(x, meta.data))
```

# Make heatmaps of abundance data: overlaps - all abundances

```{r}
overlap.scaled.data <- lapply(overlap.mf.list, function(y) apply(y, 1, function(x) log2(x/(mean(x)))))

colors <- colorRampPalette(c("white", "lightblue", "darkblue"))(50)

plots.overlap.all <- mapply(function(x,y) (pheatmap::pheatmap(t(x[order(meta.data$group),]), cluster_rows = F, cluster_cols = F, main = y, 
                                             color = colors,
                                             show_rownames = T,
                                             show_colnames = T,
                                             silent = TRUE)$gtable), 
                                           overlap.scaled.data, 
                                           names(overlap.scaled.data),
                                           SIMPLIFY = FALSE)

pdf(file = "../outputs/heatmap_overlap_feats.pdf", width = 24, height = 16)

grid.arrange(grobs = plots.overlap.all, ncol = 6)

dev.off()

mapply(function(x,y) (pheatmap::pheatmap(t(x[order(meta.data$group),]), cluster_rows = F, cluster_cols = F, main = y, 
                                             color = colors,
                                             show_rownames = F,
                                             show_colnames = F)), overlap.scaled.data, names(overlap.scaled.data))
```

# Make heatmaps of abundance data: overlaps - averaged abundances

```{r}
overlap.scaled.data.avg <- lapply(overlap.mf.avg, function(y) apply(y, 1, function(x) log2(x/(mean(x)))))

colors <- colorRampPalette(c("white", "lightblue", "darkblue"))(50)

plots.overlap.all <- mapply(function(x,y) (pheatmap::pheatmap(t(x[order(meta.data.avg$group),]), cluster_rows = F, cluster_cols = F, main = y, 
                                             color = colors,
                                             show_rownames = T,
                                             show_colnames = T,
                                             silent = TRUE)$gtable), 
                                           overlap.scaled.data.avg, 
                                           names(overlap.scaled.data.avg),
                                           SIMPLIFY = FALSE)

pdf(file = "../outputs/heatmap_overlap_feats_averaged.pdf",width = 24, height = 16)

grid.arrange(grobs = plots.overlap.all, ncol = 6)

dev.off()

mapply(function(x,y) (pheatmap::pheatmap(t(x[order(meta.data.avg$group),]), cluster_rows = F, cluster_cols = F, main = y, 
                                             color = colors,
                                             show_rownames = F,
                                             show_colnames = F)), anova.scaled.data.avg, names(anova.scaled.data.avg))
```
